### 2015-05-21 rscripts/data_review.Rmd
exploring population more

```{r, echo=FALSE}
src.dt <- fread("../input/merged.o", header = F, sep=" ", colClasses = c("integer","integer","integer","integer"))
setnames(src.dt, c("user_id", "location_id", "login", "logout"))
setkey(src.dt, login, logout, user_id, location_id)
src.dt
```

@@

1. remove sessions exceeding `r max_hours` hours
2. eliminate users that have fewer than `r min_logins` logins
3. eliminate locations that have fewer than `r min_logins` logins
4. recursively apply steps 2-3 until no users / locations eliminated

```{r, echo=FALSE}
censor.dt <- src.dt[(logout - login) <= max_hours*60*60, ]
invalid.users <- censor.dt[,.N, by=user_id ][N < min_logins, user_id]
while(length(invalid.users) > 0) {
  censor.dt <- censor.dt[!user_id %in% invalid.users]
  invalid.locs <- censor.dt[,.N, by=location_id ][N < min_logins, location_id]
  censor.dt <- censor.dt[!location_id %in% invalid.locs]
  invalid.users <- censor.dt[,.N, by=user_id ][N < min_logins, user_id]
}
censor.dt
```

@@

```{r, echo=FALSE}
source("munging.R")
breakoutDays(censor.dt[,
  user_id := .GRP, by=user_id
][,
  location_id := .GRP, by=location_id
])
setkey(censor.dt, login_day, login_time, logout_day, logout_time, user_id, location_id)
censor.dt
```

@@

>>>>

```{r, echo=FALSE}
src.dt <- 
  setkey(
    setnames(
      fread("../input/merged.o", header = F, sep=" ", colClasses = list(integer=1.4)),
      c("user_id", "location_id", "login", "logout")
    ),
    login, logout, user_id, location_id
  )
src.dt
```

@@

1. remove sessions exceeding `r max_hours` hours
2. remove sessions with no time (these correspond to rows in the source data with no logout time)
3. eliminate users that have fewer than `r min_logins` logins
4. eliminate locations that have fewer than `r min_logins` logins
5. recursively apply steps 2-3 until no users / locations eliminated

```{r, echo=FALSE}
censor.dt <- src.dt[(logout != login) & ((logout - login) <= max_hours*60*60), ]
invalid.users <- censor.dt[,.N, by=user_id ][N < min_logins, user_id]
while(length(invalid.users) > 0) {
  censor.dt <- censor.dt[!user_id %in% invalid.users]
  invalid.locs <- censor.dt[,.N, by=location_id ][N < min_logins, location_id]
  censor.dt <- censor.dt[!location_id %in% invalid.locs]
  invalid.users <- censor.dt[,.N, by=user_id ][N < min_logins, user_id]
}
censor.dt
```

@@

```{r, echo=FALSE}
source("munging.R")
setkey(breakoutDays(censor.dt[,
  user_id := .GRP, by=user_id
][,
  location_id := .GRP, by=location_id
]), login_day, login_time, logout_day, logout_time, user_id, location_id)
censor.dt
```

@@

Our synthetic population comprises long term users of the system.  When synthezing their behavior, we should be thinking relative to other long term users of the system.

```{r, echo=FALSE}
reference.dt <- merge(censor.dt, censor.dt[,list(tot = (max(logout) - min(login))/(365*24*60*60)), by=user_id][tot > 1, list(user_id = user_id)], by="user_id")
setkeyv(reference.dt, key(censor.dt))
ref.dt <- reference.dt[,
  list(
    overnight_logins = sum(login_day != logout_day),
    non_overnight_logins = sum(login_day == logout_day),
    mean_session_min = mean(logout-login)/60,
    total_session_min = sum(logout-login)/60,
    user_lifetime_days = max(logout_day) - min(login_day)
), by=user_id][,total_logins := overnight_logins + non_overnight_logins]
hist(ref.dt$overnight_logins/c(ref.dt$overnight_logins + ref.dt$non_overnight_logins))
hist(ref.dt$mean_session_min)
fitting <- function(runs, p_continue) sum(rnbinom(n=runs, size=1, prob=1-p_continue))/runs
with(ref.dt, {
  plot(total_logins, mean_session_min)
  points(total_logins, sapply(total_logins, fitting, p_continue = 0.99), pch=".", col = "red")
})

plot()
```
